---
title: "BST260 Final Project"
author: "Takahisa Ogawa\v  Kazuma Oyama\v  Keitaro Mahara\v  Keiko Kunitoki\v"
date: "12/12/2019"
output: html_document
---

## Impact of distance to hospital on amputation rate in trauma patients

##  Introduction

• The number of amputation has recently been increasing.

• Amputation largely affect the patients' quality of life.

• The success for limb salvage depends on standby time before surgery.

• Logistics such as distance from injury site to a hospital has an important role on the time before surgery

## Objectives

• To investigate the association between the distance to a hospital and amputation rate in trauma patients.

## Distance

• The U.S. Geological Survey developed the Geographic Names Information System (GNIS) for the U.S. Board on Geographic Names.

• Distance is calculated by ArcGIS.

• Used minimum distance(Meter)

## Data wrangling

```{R}
library(tidyverse)
library(readr)


data <- read_delim("injurydata.csv",";", escape_double = FALSE, trim_ws = TRUE)

#cleaning_employer 
data$EMPLOYER =gsub(".*usps|us postal|united states postal|u.s. postal|u.s postal|u. s postal|u. s. postal.*","US_Postal_Service",                   ignore.case = TRUE, data$EMPLOYER)
data$EMPLOYER =gsub(".*US_Postal_Service.*","USPS", ignore.case = TRUE, data$EMPLOYER)
data$EMPLOYER =gsub(".*united parcel|ups |ups,.*","United_Parcel_Service", ignore.case = TRUE, data$EMPLOYER)
data$EMPLOYER =gsub(".*United_Parcel_Service.*","UPS", ignore.case = TRUE, data$EMPLOYER)
data$EMPLOYER =gsub(".*american airl.*","American Airlines", ignore.case = TRUE, data$EMPLOYER)
data$EMPLOYER =gsub(".*AT &|AT&.*","AT_T", ignore.case = TRUE, data$EMPLOYER)
data$EMPLOYER =gsub(".*AT_T.*","AT&T Inc", ignore.case = TRUE, data$EMPLOYER)
data$EMPLOYER =gsub(".*walmart|wallmart|wal-mart.*","wal_mart", ignore.case = TRUE, data$EMPLOYER)
data$EMPLOYER =gsub(".*wal_mart.*","Walmart", ignore.case = TRUE, data$EMPLOYER)
data$EMPLOYER =gsub(".*Publix.*","Publix_", ignore.case = TRUE, data$EMPLOYER)
data$EMPLOYER =gsub(".*Publix_.*","Publix", ignore.case = TRUE, data$EMPLOYER)
data$EMPLOYER =gsub(".*Asplundh.*","Asplundh_", ignore.case = TRUE, data$EMPLOYER)
data$EMPLOYER =gsub(".*Asplundh_.*","Asplundh", ignore.case = TRUE, data$EMPLOYER)
data$EMPLOYER =gsub(".*sodexo.*","sodexo_", ignore.case = TRUE, data$EMPLOYER)
data$EMPLOYER =gsub(".*sodexo_.*","Sodexo", ignore.case = TRUE, data$EMPLOYER)
data$EMPLOYER =gsub(".*Waste Management.*","Waste_Management", ignore.case = TRUE, data$EMPLOYER)
data$EMPLOYER =gsub(".*Waste_Management.*","Waste Management", ignore.case = TRUE, data$EMPLOYER)
data$EMPLOYER =gsub(".*Tyson Foods.*","Tyson_Foods", ignore.case = TRUE, data$EMPLOYER)
data$EMPLOYER =gsub(".*Tyson_Foods.*","Tyson Foods", ignore.case = TRUE, data$EMPLOYER)

#add SES
#a <-read_csv("ACS_17_5YR_S0101_with_ann.csv")
#b  <-read_csv("ACS_17_5YR_S0701_with_ann.csv")
#c <-read_csv("ACS_17_5YR_S0801_with_ann.csv")
#d <-read_csv("ACS_17_5YR_S1101_with_ann.csv")
#e <-read_csv("ACS_17_5YR_S1901_with_ann.csv")
#f <-read_csv("ACS_17_5YR_S2301_with_ann.csv")
#a<-a %>%   rename('ZIP' = 'GEO.id2')
#b<-b %>%   rename('ZIP' = 'GEO.id2')
#c<-c %>%   rename('ZIP' = 'GEO.id2')
#d<-d %>%   rename('ZIP' = 'GEO.id2')
#e<-e %>%   rename('ZIP' = 'GEO.id2')
#data <- data %>% left_join(a,by="ZIP")
#data <- data %>% left_join(b,by="ZIP")
#data <- data %>% left_join(c,by="ZIP")
#data <- data %>% left_join(d,by="ZIP")
#data <- data %>% left_join(e,by="ZIP")


#add_state population
f <-read_csv("state_pop.csv",col_names = F)
f <- rename(f,STATE=X1)
f <- rename(f,POP=X2)
datax <- data %>% left_join(f, by="STATE")


#filter hospitarisation(0/1)
dataa <- datax %>% filter(HOSPITALIZED<2)
datas <- dataa %>% filter(HOSPITALIZED>-1)
#filter amputation(0/1)
datad <- datas %>% filter(AMPUTATION<2)
dataf <- datad %>% filter(AMPUTATION>-1)
#filter by hospital id
dataq <- dataf %>% filter(NEAR_FID>0)

#distance less than 10^5 meter
data <- dataq %>% filter(NEAR_DIST<10^5)

#add_multiple injury
Left<-function(x,chr){
  return(substr(x,1,chr))}

data<-mutate(data, leftMul = Left(PARTOFBODYTITLE,3))
data<-mutate(data, MultiInj = ifelse(data$leftMul=="Mul", "1", "0"))
data
#### åºå
##write.csv(datag, "C:/Users/keiko/Desktop/BST260-PROJECT/injurydata1130.csv", sep=";" ) 
```


## Relationship betweem population and the number of injury according to states

```{r}
# group by state
injury_state <- data %>% group_by(STATE) %>% summarise(number = n())

# add population data
injury_state_pop <- transform(injury_state, population = c(4898246,735720,7275070,3026412,39747267,5770545,3567871,975033,711571,29087070,10627767,162742,1416589,1790182,12700381,6718616,3167997,2910931,4484047,4652581,1342097,6062917,6939373,10020472,5655925,2987895,6147861,1074532,1940919,3087025,1363852,8922547,2096034,19491339,10497741,760900,55144,11718568,3948950,4245901,12813969,3113659,1056738,5147111,892631,6833793,29087070,3221610,627180,106405,8571946,7666343,1791951,5832661,572381)) %>% filter(number > 50)
# visualize
ggplot(data = injury_state_pop, aes(x = population, y = number, label = STATE)) +
  labs(title = "Relationship betweem population and the number of injury according to states ", x = "Population", y = "Number of injury") + geom_point()
```

```{r}
pie_chart_hosp <- data %>% select(HOSPITALIZED)
table(pie_chart_hosp)
pie_chart_hosp2 <- c(8881,36041)
pie(pie_chart_hosp2, radius=1, labels=c("Non-Hospitalzed 19.8%","Hospitalized 80.2%"), col=c("#0000FF","#FF8000"), main="Hospitalization Rate in All Injuries")
```

```{r}
pie_chart_amp1 <- data %>% select(AMPUTATION)
table(pie_chart_amp1)
pie_chart_amp2 <- c(33173,11997)
pie(pie_chart_amp2, radius=1, labels=c("Non-Amputation 73.4%","Amputation 26.6%"), col=c("#BFFF00","#FF0000"), main="Amputation Rate in All Injuries")
```

```{r}
data_ah_rate <- data %>% group_by(STATE) %>%
  summarise(amputation_rate = mean(AMPUTATION),   hospitalazation_rate = mean(HOSPITALIZED), 
            number = n()) %>%
  filter(number > 50)

ggplot(data = data_ah_rate, aes(x = hospitalazation_rate, y = amputation_rate, label = STATE)) +
  labs(title = "Relationship betweem Hospitalization and Amputation Rates according to States ", x = "Hospitalization Rate", y = "Amputation Rate") + geom_point() 

```

## longitudinal analysis 
```{r}
library(lubridate)
data2 <- data %>% mutate(month = month(EVENTDATE), year = year(EVENTDATE)) 
timeseries <- data2%>% group_by(year,month) %>%
  summarise(amputation =sum(AMPUTATION)) %>% inner_join(data2, by=c('year'='year','month'='month'))

timeseries <- data2 %>% group_by(year, month) %>%
  mutate(amputation =sum(AMPUTATION)) 

timeseries %>% ggplot()+aes(EVENTDATE,amputation) +geom_line()+
  xlab('Date')+ylab('Number of Amputaions (people)')

timeseries %>% ggplot()+aes(EVENTDATE,amputation) +geom_smooth()+
  xlab('Date')+ylab('Number of Amputaions (people)')
```

## Natural Language Processing Infrastructure (NLP)

```{r}
library(tidytext)
library(tidyverse)
library(wordcloud2)

#NLP
text <- data %>%  unnest_tokens(word, FINALNARRATIVE)
data(stop_words)
text <- text %>% anti_join(stop_words)
text <-transform(text,Freq=0)

#Bar chart
text %>%
  count(word, sort = TRUE) %>%
  filter(n > 5000) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n)) +
  geom_col() +
  xlab(NULL) +
  coord_flip()

#Word Cloud
library(wordcloud2)
wordcloud2(text,size=1,minSize=10)
wctext <- text %>% group_by(word) %>% mutate(Freq = n())

wctext%>% select(word,Freq)%>% 
  distinct(word, .keep_all = TRUE)%>%
  arrange(desc(Freq)) %>% 
  wordcloud2()

```

## Distance from hospital according to states

```{r distance and states, echo=FALSE}

# group by states and summarise
data_bar_s <- data %>% group_by(STATE) %>%
  summarise(mean = mean(NEAR_DIST), sd =sd(NEAR_DIST), se = sd(NEAR_DIST)/sqrt(n())) 

# visualize
data_bar_s %>% mutate(STATE = reorder(STATE, mean)) %>%
  ggplot(aes(x = STATE, y = mean, fill= STATE)) +
  geom_bar(stat = "identity")+
  geom_errorbar(aes(ymin=mean - se, ymax= mean +se)) +
  theme(legend.position = 'none') +
  xlab("State")+
  ylab("distance from hospital (m)") +
   coord_flip() 

```

## Amputation rate and distance according to states on us map

```{r}
# group by states and summarise
data_amputation_rate <- data %>% group_by(STATE) %>%
  summarise(mean = mean(NEAR_DIST), sd =sd(NEAR_DIST), se = sd(NEAR_DIST)/sqrt(n()),amputation_rate = mean(AMPUTATION), amputation =sum(AMPUTATION)) 

# manipulate data for the map
# use us map data and lefh_join by region
data_amputation_rate_1 <- data_amputation_rate %>%
  mutate(region = str_to_lower(STATE))
library(maps)
states_map <- map_data("state")
amputation_map <- left_join(states_map, data_amputation_rate_1, by = "region")

# Create the map
library(ggthemes)
amputation_map %>%
ggplot(mapping = aes(long, lat, group = group,fill = amputation_rate))+
  geom_polygon(color = "black", size = 0.1) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(title = "Amputation Rate") + 
  theme_map() +
  scale_fill_gradient(low = "white", high = "#CB454A")
```

## Amputation rate and distance according to states
```{r}
# visualize on scatterplot
data_amputation_rate <- data_amputation_rate %>% filter(amputation >= 10) 
  data_amputation_rate  %>%
  ggplot(aes(x = mean, y = amputation_rate, label = STATE)) +
  geom_point()+
  xlab("Distance from hospital (m)")+
  ylab("Amputation rate")+
  ggtitle("Amputation Rate and Distance by State")

# add state abbreviation and change size according to the number of amputation
library(openintro)
library(ggrepel)

data_amputation_rate %>% 
  mutate(abb = state2abbr(STATE)) %>%
  ggplot(aes(x = mean, y = amputation_rate, label = abb, size = amputation)) +
  geom_point()+
  geom_text_repel(size =3) +
  xlab("Distance from hospital(m)")+
  ylab("Amputation rate") +
  xlim(0,15000) +
  ggtitle("Amputation Rate and Distance by State")
# add regression line
data_amputation_rate %>% 
  ggplot(aes(x = mean, y = amputation_rate, label = STATE)) +
  geom_point()+
  geom_smooth(method="lm") +
  xlab("Distance from hospital(m)")+
  ylab("Amputation rate") +
  xlim(0,15000) +
 ggtitle("Amputation Rate and Distance by State")

# check correlation
tidy(cor.test(data_amputation_rate$mean, data_amputation_rate$amputation_rate))

```

## Hospital location on injury map

```{r}
library(maps)
library(ggthemes)

# make us map
us_map <- ggplot()+
  geom_polygon(data = states_map, aes(long, lat, group = group), color = "black", size = 0.1, fill = "white") +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  theme_map() 
us_map

# add injury data on us map
data_map <- data %>% mutate(LONGITUDE = as.numeric(LONGITUDE), LATITUDE =as.numeric(LATITUDE), AMPUTATION = as.factor(AMPUTATION)) %>% filter(LATITUDE > 20 & LATITUDE < 50 & LONGITUDE > -120)

injury_map <- us_map + geom_point(data = data_map, aes(x=LONGITUDE, y=LATITUDE, color = AMPUTATION), size=0.0001) +
  ggtitle("Injury and Amputation Map")
injury_map

# add hospital location on injury map
hosp_data <- read_csv("HPDATA.csv") %>% mutate(long= as.numeric(LONGI), lat= as.numeric(LATIT)) %>% 
  filter(lat> 20 & lat< 50 & long > -120)
injury_map + geom_point(data = hosp_data, aes(x=long, y=lat), color = "black", size = 0.1, alpha = 0.5) +
  ggtitle("Injury and Hospital Map")


```




## bar graph
```{r amputaion and distance}

data_bar <- data %>% mutate(AMPUTATION = as.factor(AMPUTATION)) %>%
                            group_by(AMPUTATION) %>%
  summarise(mean = mean(NEAR_DIST), sd =sd(NEAR_DIST), se = sd(NEAR_DIST)/sqrt(n())) 
data_bar 

data_bar %>%
  ggplot(aes(x = AMPUTATION, y = mean, fill=AMPUTATION)) +
  geom_bar(stat = "identity", width = 0.5)+
  geom_errorbar(aes(ymin=mean - se, ymax= mean +se), width = 0.3) +
  xlab("amputation")+
  ylab("distance from hospital (m)")
```



## boxplot
This graph compares the distance from hospital to the injury place between those who took amputation and those who didn't. Please note the y axis is described in log 10 scale in meter.
```{r}
data$HOSPITALIZED <- as.numeric(data$HOSPITALIZED)
data$MultiInj <- as.factor(data$MultiInj)
data$AMPUTATION <- as.factor(data$AMPUTATION)
data %>%
  ggplot(aes(x=AMPUTATION,y=log10(NEAR_DIST)))+
  geom_boxplot()+
  xlab("amputation")+
  ylab("distance from hospital log10")
```
The median distance of those who took amputation seems longer. 

## density graph
This graph also comparing the distance from hospital between those who took amputation and not. The peak of data are in 1,000 to 10,000 meter. 

blue: Amputation +   red: amputation -
```{r}
data %>% ggplot(aes(log10(NEAR_DIST), fill=AMPUTATION)) + geom_density(alpha = 0.2)+
  xlab("distance to hospital, log10, meter")
```
## ttest
Then, we performed stutent t test. There is a significant difference in distance from hospital(p<0.001).

```{r}
t.test(data$NEAR_DIST~data$AMPUTATION)
```
Mean distance: Amp- 5860(m)      Amp+ 6267 (m)

## regression
### univariate Regression　(distance)
First, we saw the crude relationship between distance from hospital and probability of amputation.
```{r}
mlog <- glm(AMPUTATION ~ NEAR_DIST , data = data, family = "binomial")
summary(mlog)
```
### Odds Ratio
```{r}
exp(cbind(OR = coef(mlog), confint(mlog)))
```
#### probability graph for univariate logistic regression
```{r}
#create values
fit_1 <- with(data, data.frame(NEAR_DIST = rep(seq(from = 1, to = 100000, length.out = 1000), 2)))
#generate
fit_2 <- cbind(fit_1, predict(mlog, newdata = fit_1, type = "link",se = TRUE))
fit_2 <- within(fit_2, {
    PredictedProb <- plogis(fit)
    LL <- plogis(fit - (1.96 * se.fit))
    UL <- plogis(fit + (1.96 * se.fit))
})
## view first few rows of final dataset
head(fit_2)
nrow(fit_2)
#plot
ggplot(fit_2, aes(x = NEAR_DIST, y = PredictedProb)) + geom_ribbon(aes(ymin = LL,
    ymax = UL), alpha = 0.2) + geom_line()+xlab("Distance from Hospital")+ylab("Predicted Probability of Amputation")+xlim(100,50000)
ggplot(fit_2, aes(x = log10(NEAR_DIST), y = PredictedProb)) + geom_ribbon(aes(ymin = LL,
    ymax = UL), alpha = 0.2) + geom_line()+xlab("Distance from Hospital")+ylab("Predicted Probability of Amputation")
```
### multivariate logistic Regression　(distance, hospitalization, multiple injury+/-)
```{r}
mylogit <- glm(AMPUTATION ~ NEAR_DIST  + HOSPITALIZED + MultiInj, data = data, family = "binomial")
summary(mylogit)
```
#### Odds Ratio
```{r}
exp(cbind(OR = coef(mylogit), confint(mylogit)))
```
Estimated Odds Ratio for 1 m increase in distance is 1.000008 (95% CI: 1.000003-1.000013).

As a result, 1 km increase in distance from hospital raise 1.008 times Odds of amputation, adjusting hospitalization and multiple injury.

## conclusion

• According to prior knowledge insisting time to hospital is related to outcome of injury, there is a significant relationship between distance from hospital and amputation rate.
• The US, the country having vast land, have to care about the disposition of hospitals.
